\vspace*{70mm}
\begin{center}
    \renewcommand{\baselinestretch}{1.0}
    \sl
    To \ldots
\end{center}

\chapter{Acknowledgements}

My utmost respect and appreciation must be stated for those individuals who, whilst experiencing some of the toughest times of their lives in homelessness, were willing to participate in this project. Alongside these individuals I also wish to voice my appreciation for the staff, volunteer, and donor communities in the Not-for-Profits surrounding these individuals with much-needed support and service.

I would like to share my deepest appreciations for Dr Dhaval Vyas, whom without this project would not have been possible. Beyond his initial investment in this project, Dr Vyas has enthusiastically and devotedly supported this project throughout its many trials and tribulations. I am incredibly grateful for the expertise and patience that has been provided by Dr Vyas, particularly in navigating the unprecedented global landscape of 2020 and all of its accompanying concerns. In times of diminishing motivation and morale, Dr Vyas always remained positive, encouraging, and optimistic, providing a guiding light towards this project's completion.

The commencement and undertaking of this project would also not have been possible without the passion and camaraderie provided by my colleagues, Jay Almaraz and Stephen Pozzi. Not only were these two crucial in the initial designs of this project, but they have proven to be consistently motivating as we all strive to complete our own individual works. Jay and Stephen also played a pivotal role in data collection, delivery logistics, and communication channels in their altruistic support of this project.

To Orange Sky, who enabled this otherwise very academically minded project to ground itself in reality, I am forever grateful. Without Orange Sky, whom I spent many of my most formative years with, I would not have been in a position to undertake this project. I am indebted to you for the growth, both personally and professionally, which I underwent during my time on the team.

On a personal note, the unwavering support from my partner Kelsie and my furry companion Sienna, have driven and motivated me beyond reach of anything I could have achieved myself. I look forward to the extra time that completing this project will allow us to spend together.

I also must acknowledge my parents, one more academically empathetic than the other, for reasons beyond itemisation in a format as short as this. 

\cleardoublepage